Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-05-11 paper/main.tex
related work, measurment scale, bundles

\subsection{Equivariance}
\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{equiviz.png}
\end{figure}

@@

\begin{figure}[h]
  \label{fig:atct:fb}
  \includegraphics[width=\columnwidth]{fiberbundle.png}
  \caption{change to concrete example}
\end{figure}

@@

\begin{figure}
  \label{fig:atct:fb:graphic}
  \includegraphics[width=1\columnwidth]{render.png}
  \caption{}
\end{figure}
In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of the \note{pin down what's beeing drawn}.   
The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber. The preimage of the section at these coordinates returns a set of coordinates ${\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}$ at the points $\gbasepoint \in \gbase$ in the graphic base topological space. The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber. Each point has an RGB tuple and these tuples are integrated to return the color of the pixel at that $(x,y)$ point.

@@

In this work, we assume that the map
\begin{equation}
 \label{eq:atct:morphism:base}
\dfunch: \openset^{\prime}\rightarrow \openset
\end{equation}
is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$. For example a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values. This base space transformation induces a pullback section 
\begin{equation}
  \label{eq:atct:morphism:basepull}
  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
\end{equation}
such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$. This means that the base space transformation \dfunch\ alone does not change the data values.

As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Examples of a fiber transformation are actions on fibers, such as the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.

The full transformation 
\begin{equation}
  \label{eq:atct:morphism:all}
  \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
\end{equation}
is used to express actions such as rotation that induce both a remapping of the index space and a change in the data values. The data transform \dfunc\ is composable
\begin{equation}
  \dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)
\end{equation}
if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ such that $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$. This allows us to define a data transform where each fiber transform $\dfunct_{i}$ can be applied to a different fiber field $\dfiber_i$.

@@

\begin{figure}[ht]
  \label{fig:atct:morphisms:sheaf}
  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
  \caption{}
\end{figure}

>>>>

\subsection{Equivariance}
\begin{figure}[h!]
  \label{fig:related-work:equivariance}
  \includegraphics[width=1\columnwidth]{equiviz.png}
\end{figure}

@@

\begin{figure}[h!]
  \label{fig:atct:fb}
  \includegraphics[width=\columnwidth]{fb_tau.png}
  \caption{change to concrete example}
\end{figure}

@@

\begin{figure}[h!]
  \label{fig:atct:fb:graphic}
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{}
\end{figure}
In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of the \note{pin down what's beeing drawn}.   
The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber. The preimage of the section at these coordinates returns a set of coordinates ${\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}$ at the points $\gbasepoint \in \gbase$ in the graphic base topological space. The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber. Each point has an RGB tuple and these tuples are integrated to return the color of the pixel at that $(x,y)$ point.

@@

In this work, we assume that the map
\begin{equation}
 \label{eq:atct:morphism:base}
\dfunch: \openset^{\prime}\rightarrow \openset
\end{equation}
is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$. For example a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values. This base space transformation induces a pullback section 
\begin{equation}
  \label{eq:atct:morphism:basepull}
  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
\end{equation}
such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$. This means that the base space transformation \dfunch\ does not change the data values at a given point
\begin{equation}
  \label{eq:atct:morphism:verify_base}
  \dsection(\dbase) = \dfuncpull\dsectionc(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbase^{prime}))
\end{equation} 
where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$. As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.

We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
\begin{equation}
  \label{eq:atct:morphism:all}
  \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
\end{equation}
which gives us an equation that can express complex transformations. For example, wind vectors on a sphere are a non-trivial bundle, meaning their fibers aren't identical at every point in the base space; therefore a rotation (change of base space) of the sphere must include an adjustment to the values that have moved closer to the pole (a change in fiber).  
The data transform \dfunc\ is composable 
\begin{equation}
  \dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)
\end{equation}
if each (identical) component base space is transformed in the same way $\dfunch$ and there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ such that $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$. This allows us to define a data transform where each fiber transform $\dfunct_{i}$ can be applied to a different fiber field $\dfiber_i$.

\begin{figure}[h!]
  \label{fig:atct:phi}
  \includegraphics[width=\columnwidth]{phi.png}
\end{figure}
Transposition, as illustrated in \autoref{fig:atct:phi}, is one example of a transformation on the base space $(\dfunch, \dfuncpull)$; transposition is a continuity preserving data transformation that moves every data value to a new location such that the values are unchanged and connected in the same way. Scaling, as shown in \autoref{fig:atct:phi} is a transform $\dfunct$ that changes each data value, in this example by multiplying each value by two. As defined in \autoref{eq:atct:morphism:all}, these transformations can combine to yield a matrix that is both transposed and scaled. The matrix in \autoref{fig:atct:phi} is visualized as a heatmap \autoref{fig:related-work:equivariance} and, as discussed in \autoref{sec:related-work:equivariance}, the heatmap is illustrating equivariance. In the $(\dfunch, \dfuncpull)$ case, the data to color map has not changed, but the column bands of color have become row bands. In the $\dfunct$ case, the data has been scaled by two and the color encoding has been scaled by two, but the bands remain column ordered. When the data is transposed and scaled, then the heatmap is also transposed and the colors are scaled.

@@

\begin{figure}[h!]
  \label{fig:atct:morphisms:sheaf}
  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
  \caption{}
\end{figure}
